---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
This is an R markdown File of Figures presented in ""

Libraries 
```{r message=FALSE}
library(UpSetR)
library(sna)
library(plotrix) 
library(ggplot2)
library(clusterProfiler)
library(org.Hs.eg.db)
```

```{r echo=FALSE}
# interaction plot function 
intersection.plot <- function (nameList, dbList, compNameList, compList){
  l <- length(nameList)
  cl <- length(compNameList)
  
  # add totals under names 
  for(i in 1:l){
    nameList[i] <- paste0(paste0(nameList[i],"\n("),toString(length(dbList[[i]])),")")
  }
  for(i in 1:cl){
    compNameList[i] <- paste0(paste0(compNameList[i],"\n("),toString(length(compList[[i]])),")")
  }
  
  # calculate numbers and percent 
  Final <- matrix(0, ncol = l, nrow = cl)
  Final2 <- matrix(0, ncol = l, nrow = cl)
  for(i in 1:l){
    for (j in 1:cl){
      Final2[j,i] <- (length(intersect( dbList[[i]], compList[[j]]))) # numbers 
      Final[j,i] <- (length(intersect(  dbList[[i]], compList[[j]])))/length(compList[[j]]) #percent 
    }
  }
  
  Final <- as.data.frame(Final)
  colnames(Final) <- nameList
  rownames(Final) <- compNameList
  
  Final2 <- data.matrix(Final2)
  colnames(Final2) <- nameList
  rownames(Final2) <- compNameList
  
  
  Final <- as.matrix(Final)
  mode(Final2) <- "integer"
  
  cellcol<-color.scale(Final, extremes = c("white", "Black")) # intersect greyscale
  
  par(mar=c(1,10,4,1))
  color2D.matplot(Final2, show.values = .1 , axes= FALSE , cellcolors = cellcol, xlab = "", ylab = "", border = "black", vcex = 1.6 )
  axis(3,at=0.5:l, las= 0, labels=colnames(Final),cex.axis = 1, family= "sans")
  axis(2,at=0.5:cl,las=2,labels=rev(rownames(Final)),cex.axis = 1, family = "sans")
}

# interaction plot function 
setDifference.plot <- function (nameList, dbList, compNameList, compList){
  l <- length(nameList)
  cl <- length(compNameList)

    # add totals under names 
  for(i in 1:l){
    nameList[i] <- paste0(paste0(nameList[i],"\n("),toString(length(dbList[[i]])),")")
  }
  for(i in 1:cl){
    compNameList[i] <- paste0(paste0(compNameList[i],"\n("),toString(length(compList[[i]])),")")
  }
  
  # calculate numbers and percent 
  Final <- matrix(0, ncol = l, nrow = cl)
  Final2 <- matrix(0, ncol = l, nrow = cl)
  for(i in 1:l){
    for (j in 1:cl){
      Final2[j,i] <- (length(setdiff( dbList[[i]], compList[[j]]))) # numbers 
      Final[j,i] <- (length(setdiff(  dbList[[i]], compList[[j]])))/length(compList[[j]]) #percent 
    }
  }
  
  Final <- as.data.frame(Final)
  colnames(Final) <- nameList
  rownames(Final) <- compNameList
  
  Final2 <- data.matrix(Final2)
  colnames(Final2) <- nameList
  rownames(Final2) <- compNameList
  
  
  Final <- as.matrix(Final)
  mode(Final2) <- "integer"
  
  cellcol<-color.scale(Final, extremes = c("white", "Black")) # intersect greyscale
  
  par(mar=c(1,10,9,1))
  color2D.matplot(Final2, show.values = .1 , axes= FALSE , cellcolors = cellcol, xlab = "", ylab = "", border = "black", vcex = 1.6 )
  axis(3,at=0.5:l, las= 0, labels=colnames(Final),cex.axis = 1, family= "sans")
  axis(2,at=0.5:cl,las=2,labels=rev(rownames(Final)),cex.axis = 1, family = "sans")
  title("Set Difference")
}

erichment.plot <- function (uniqueList, background, title){
  ego <- clusterProfiler::enrichGO(
    gene = uniqueList,      # a vector of UniProt ids
    keyType = "UNIPROT",
    universe = ultimate,
    OrgDb = "org.Hs.eg.db",      # organism to analyze
    ont = as.character('MF'),    # ontology term ("BP", "MF", "CC")
    pAdjustMethod = "BH",        # method for multiple testing correction 
    pvalueCutoff = 0.05,         # P-value cutoff 
    qvalueCutoff = 0.05,         # default to 0.2
    readable = TRUE,            # return gene symbols
    minGSSize = 5,              # minimal size of genes annotated by Ontology term for testing.
    maxGSSize = 10000
  )
  p <- barplot(ego, showCategory = 40, title = title)
  p
}
```

```{r echo=FALSE}
dn <- "./"

# Proteins
PSP <- read.csv( paste0(dn,"PSP/UniProtIDs_noISO.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
RXM <- read.csv(paste0(dn, "RXM/UniProtIDs_noISO.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
HPRD <- read.csv(paste0(dn, "HPRD/UniProtIDs_noISO.tsv" ), header = FALSE, stringsAsFactors = FALSE)[,1]
BG <- read.csv(paste0(dn, "BG/UniProtIDs_noISO.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
KEGG <- read.csv(paste0(dn, "KEGG/updatedAllUID.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
WP <- read.csv(paste0(dn, "WP/updatedAllUID.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
SIG <- read.csv(paste0(dn, "SIGNOR/UniProtIDs_noISO.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]


# Phosphosites
PSP_mods <- read.csv(paste0(dn, "PSP/Phosphorylations_NoISO.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
RXM_mods <- read.csv(paste0(dn, "RXM/Phosphorylations_NoISO.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
HPRD_mods <- read.csv(paste0(dn,"HPRD/Phosphorylations_NoISO.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
BG_mods <- read.csv(paste0(dn,"BG/Phosphorylations_NoISO.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
UNIPROT_mods <- read.csv(paste0(dn, "UNIPROT/SwissProtMods.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
SIG_mods <- read.csv(paste0(dn, "SIGNOR/Phosphorylations_NoIso.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]

# reference databases 
QPHOS <- read.csv(paste0(dn,"QPHOS/updatedAllUID_qphos.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
IMEX <- read.csv(paste0(dn,"IMEX/updatedAllUID.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
UNIPROT <- read.csv("UIDs.tsv", header = FALSE, stringsAsFactors = FALSE)[,1]
GO_SIG <- read.csv(paste0(dn,"GO/updatedAllUID_SIG.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_mods <- read.csv(paste0(dn, "QPHOS/updatedAllUID_qphos_mods.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
GO_PKA <- read.csv( paste0(dn, "GO/updatedAllUID_PKA.tsv"),sep = "\t", header = TRUE, stringsAsFactors = FALSE)[,1]
UP_PP <- read.csv(paste0(dn,"UNIPROT/UP_PP.tsv"),sep = "\t", header = TRUE, stringsAsFactors = FALSE)[,1]



# Experimental data 
QPhos_HeLa_S3 <- read.csv(paste0(dn,"QPHOS/UID_HeLa_S3_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_HeLa <- read.csv(paste0(dn,"QPHOS/UID_HeLa_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_MCF7 <- read.csv(paste0(dn,"QPHOS/UID_MCF-7_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_Jurkat <- read.csv(paste0(dn,"QPHOS/UID_Jurkat_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_HEK293 <- read.csv(paste0(dn, "QPHOS/UID_HEK293_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_MCF10 <- read.csv(paste0(dn, "QPHOS/UID_MCF-10A_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]

QPhos_HeLa_S3_mods <- read.csv(paste0(dn, "QPHOS/UID_MOD_HeLa_S3_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_HeLa_mods <- read.csv(paste0(dn,"QPHOS/UID_MOD_HeLa_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_MCF7_mods <- read.csv(paste0(dn, "QPHOS/UID_MOD_MCF-7_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_Jurkat_mods <- read.csv(paste0(dn, "QPHOS/UID_MOD_Jurkat_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_HEK293_mods <- read.csv(paste0(dn, "QPHOS/UID_MOD_HEK293_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]
QPhos_MCF10_mods <- read.csv(paste0(dn, "QPHOS/UID_MOD_MCF-10A_NoIso.txt"), header = FALSE, stringsAsFactors = FALSE)[,1]

# reference UIDs
Trembl <- read.csv(paste0(dn, "UNIPROT/Trembl.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
SwissProt <- read.csv(paste0(dn, "UNIPROT/SwissProt.tsv"), header = FALSE, stringsAsFactors = FALSE)[,1]
```

## Protein Level 
UpSet Plot 
```{r fig.width=18}

listInput <- list(PhosphoSitePlus= PSP, HPRD = HPRD, BioGRID = BG, Reactome = RXM, KEGG= KEGG, WikiPathways = WP, SIGNOR = SIG)
matrix <- UpSetR::fromList(listInput)

upset(matrix,
      empty.intersections = "on",
      order.by = "freq",
      nsets = 7,
      text.scale = c(1.5, 1.5, 0, 1.5, 2, 2),
      nintersects = 20,
      scale.sets = "identity",
      sets.bar.color = c("#909cc2", "#fe7f2d","#233d4d","#a1c181", "#619b8a", "#e099dc", "#fcca46")
      )
```

Bar Plot 
```{r fig.width=12}
# specifing outside order
level_order= c("BioGRID", "Reactome","HPRD",  "KEGG", "WikiPathways","SIGNOR", "PhosphoSitePlus" )
it_len = length(level_order)

# making empty lists
db2 <- rep(0,it_len^2)
db1 <- rep(0,it_len^2)
intersects <- rep(0,it_len^2)
order <- rep(0,it_len^2)

# filling in lists
count = 0
for(i in 1:it_len){
  for(j in 1:it_len){
    count = count +1
    db1[count] <- level_order[i]
    db2[count] <-level_order[j]
    intersects[count] <- length(intersect(listInput[[level_order[i]]],listInput[[level_order[j]]] ))
    order[count] <- j
  }
}

# making and ordering df
mydf <- data.frame(db1, db2, intersects, order)
mydf$intersects <- as.numeric(mydf$intersects)

#adding in overtop bar
count = 0
for(i in 1:it_len){
  for(j in 1:it_len){
    count = count +1
    db1[count] <- level_order[i]
    db2[count] <-level_order[i]
    intersects[count] <- length(intersect(listInput[[level_order[i]]],listInput[[level_order[i]]] ))
    order[count] <- j
  }
}
A <- data.frame(db1, db2, intersects, order)


#removing extra
mydf <- rbind(A, mydf)
mydf <- mydf[order(mydf$order, decreasing = F),]


p <- ggplot(data = mydf, aes( x= factor(db1, level = level_order)  , intersects, fill = db2)) +
  geom_bar(aes(group = order), stat = "identity", position = position_dodge())+
  coord_flip()+
  scale_fill_manual(values=c("#909cc2","#233d4d","#a1c181","#fcca46","#fe7f2d","#e099dc","#619b8a" ))+
  #scale_fill_manual(values = wes_palette("Cavalcanti1"))+
  xlab("Databases") + ylab("UniProt id overlap") +
  theme_bw()
p

#hprd, bg, wp, rxm, psp, sig, kg ??

```

Intersection heatmap

```{r fig.width=12}
dbNames = list( "BioGRID","Reactome","HPRD","KEGG", "WikiPathways", "SIGNOR", "PhosphoSitePlus" )
dbList = list( BG, RXM, HPRD, KEGG, WP, SIG,PSP)
refNameList = list("qPhos","IMEX","GO Protein\nKinase Activity", "GO Signalling", "UniProt keyword:\nPhosphoProtein")
refList = list(QPHOS, IMEX, GO_PKA, GO_SIG, UP_PP)

intersection.plot(dbNames, dbList, refNameList, refList)
```

## Phosphosite Level
Upset Plot
```{r fig.width= 16}
listInput <- list(PhosphoSitePlus= PSP_mods, HPRD = HPRD_mods, Reactome = RXM_mods,  BioGRID = BG_mods, SIGNOR = SIG_mods )


upset(fromList(listInput), 
      nsets = 5,
      order.by = "freq",   
      empty.intersections = TRUE,
      text.scale = c(1.5, 1.5, 1.5, 1.5, 2, 2),
      sets.bar.color = c( "#233d4d","#fcca46","#e099dc","#909cc2", "#fe7f2d")
      )
```
Bar Plot 
```{r fig.width=12}
#specifing outside order 
level_order= c( "HPRD","PhosphoSitePlus", "SIGNOR","BioGRID","Reactome")
it_len = length(level_order)

# making empty lists 
db2 <- rep(0,it_len^2)
db1 <- rep(0,it_len^2)
intersects <- rep(0,it_len^2)
order <- rep(0,it_len^2)

# filling in lists 
count = 0
for(i in 1:it_len){
  for(j in 1:it_len){
    count = count +1
    db1[count] <- level_order[i]
    db2[count] <-level_order[j]
    intersects[count] <- length(intersect(listInput[[level_order[i]]],listInput[[level_order[j]]] ))
    order[count] <- j
  }
}

# making and ordering df
mydf <- data.frame(db1, db2, intersects, order)
mydf$intersects <- as.numeric(mydf$intersects)

#adding in overtop bar 
count = 0
for(i in 1:it_len){
  for(j in 1:it_len){
    count = count +1
    db1[count] <- level_order[i]
    db2[count] <-level_order[i]
    intersects[count] <- length(intersect(listInput[[level_order[i]]],listInput[[level_order[i]]] ))
    order[count] <- j
  }
}
A <- data.frame(db1, db2, intersects, order)


#removing extra 
mydf <- rbind(A, mydf)
mydf <- mydf[order(mydf$order, decreasing = F),]

p <- ggplot(data = mydf, aes( x= factor(db1, level = level_order)  , intersects, fill = db2, group = order)) + 
  geom_bar(aes(group = order), stat = "identity", position = position_dodge())+
  coord_flip() +
  scale_fill_manual(values=c("#909cc2","#233d4d", "#fcca46", "#fe7f2d" , "#e099dc" ))+
  xlab("Databases") + ylab("Phosphorylation overlap") +
  theme_bw()

p
#hprd, bg, psp, rxm. sig??
```

Intersection heatmap 


```{r fig.width=10}
dbNames = list( "HPRD",  "PhosphoSitePlus","SIGNOR", "BioGRID", "Reactome")
dbList = list( HPRD_mods,  PSP_mods, SIG_mods, BG_mods, RXM_mods )
refNames = list("qPhos", "SwissProt" )
refList = list(QPhos_mods, UNIPROT_mods)

intersection.plot(dbNames, dbList, refNames, refList)

```

## Experimental Data 
### Protein Level 

```{r fig.width=12}
dbNames = list( "BioGRID", "Reactome","HPRD", "KEGG", "WikiPathways", "SIGNOR","PhosphoSitePlus")
dbList = list(BG, RXM,HPRD, KEGG, WP, SIG, PSP )
qPhosNames = list("qPhos HeLa","qPhos HEK293",  "qPhos HeLa S3",  "qPhos Jurkat", "qPhos MCF-7", "qPhos MCF-10A")
qPhosList = list(QPhos_HeLa, QPhos_HEK293, QPhos_HeLa_S3, QPhos_Jurkat,  QPhos_MCF7, QPhos_MCF10 )

intersection.plot(dbNames, dbList, qPhosNames, qPhosList)
```

### Phosphosite Level 

```{r fig.width=12}
dbNames = list( "BioGRID", "Reactome","HPRD", "SIGNOR", "PhosphoSitePlus")
dbList = list(BG_mods,RXM_mods, HPRD_mods,SIG_mods, PSP_mods)
qPhosNames = list("qPhos HeLa","qPhos HEK293",  "qPhos HeLa S3",  "qPhos Jurkat", "qPhos MCF-7", "qPhos MCF-10A")
qPhosList = list(QPhos_HeLa_mods, QPhos_HEK293_mods, QPhos_HeLa_S3_mods, QPhos_Jurkat_mods,  QPhos_MCF7_mods, QPhos_MCF10_mods )

intersection.plot(dbNames, dbList, qPhosNames, qPhosList)
```

## Supplementary Figures
## Protein Level 
full UpSet Plot 
```{r fig.width=18}

listInput <- list(PhosphoSitePlus= PSP, HPRD = HPRD, BioGRID = BG, Reactome = RXM, KEGG= KEGG, WikiPathways = WP, SIGNOR = SIG)
matrix <- UpSetR::fromList(listInput)

upset(matrix,
      empty.intersections = "on",
      order.by = "freq",
      nsets = 7,
      text.scale = c(1.5, 1.5, 0, 1.5, 2, 0),
      nintersects = 1000,
      scale.sets = "identity",
      sets.bar.color = c("#909cc2", "#fe7f2d","#233d4d","#a1c181", "#619b8a", "#e099dc", "#fcca46")
      )
```

The coverage of SwissProt and Trembl by the signallign databases
```{r fig.width=12}
nameList = list( "HPRD","BioGRID","Reactome","KEGG", "WikiPathways", "PhosphoSitePlus" )
dbList = list( HPRD, BG, RXM, KEGG, WP, PSP)
compNameList = list( "SwissProt", "Trembl")
compList = list( SwissProt, Trembl)

intersection.plot(nameList, dbList, compNameList, compList)
```


```{r fig.width=12}
dbNames = list( "HPRD","BioGRID","Reactome","KEGG", "WikiPathways", "PhosphoSitePlus", "SIGNOR" )
dbList = list(HPRD,BG,RXM,KEGG, WP, PSP, SIG)
refNameList = list("qPhos","IMEX","GO Protein\nKinase Activity", "GO Signalling", "UniProt keyword:\nPhosphoProtein")
refList = list(QPHOS, IMEX, GO_PKA, GO_SIG, UP_PP)
```

### Protein level
The set difference of the signalling databases and the reference datasets. 

Numbers indicate proteins in the signalling databases that are **not** in the reference datasets.

Shade of each cell indicates the proportion of the signalling database not in the reference dataset. 
```{r fig.width=12}
setDifference.plot(dbNames, dbList, refNameList, refList)
```

The set difference of the signalling databases and the reference datasets. 

Numbers indicate proteins in the reference datasets that are **not** in the Signalling databases.

Shade of each cell indicates the proportion of the reference dataset not in the signalling database. 
```{r fig.width=12}
setDifference.plot( refNameList, refList, dbNames, dbList)
```

```{r}
dbNames = list( "HPRD", "PhosphoSitePlus","BioGRID", "Reactome", "SIGNOR")
dbList = list( HPRD_mods, PSP_mods,BG_mods ,RXM_mods, SIG_mods )
refNames = list("QPhos", "UniProt" )
refList = list(QPhos_mods, UNIPROT_mods)
```

### Phosphosite level
The set difference of the signalling databases and the reference datasets. 

Numbers indicate phosphosites in the signalling databases that are **not** in the reference datasets.

Shade of each cell indicates the proportion of the signalling database not in the reference dataset. 
```{r fig.width=12}
setDifference.plot(dbNames, dbList, refNames, refList)
```

The set difference of the signalling databases and the reference datasets. 

Numbers indicate phosphosites in the reference datasets that are **not** in the Signalling databases.

Shade of each cell indicates the proportion of the reference dataset not in the signalling database. 
```{r fig.width=12}
setDifference.plot( refNames, refList, dbNames, dbList)
```

### Over Representation Analysis of Databases
Gene Ontology Molecular Function 
```{r, fig.width=20}
ultimate <- c(PSP,RXM,HPRD,KEGG,WP, BG, SIG)
ultimate <- as.vector(unique(ultimate))
length(unique(ultimate))
```

Reactome
```{r, fig.width=20}
comp <- c(PSP,HPRD,KEGG,WP,BG, SIG)
comp <- unique(comp)
rxm_unique<- setdiff(RXM, comp)

erichment.plot(rxm_unique, ultimate,"Reactome")
```
HPRD
```{r fig.width=12}
comp <- c(PSP, KEGG,WP, RXM, BG, SIG)
comp <- unique(comp)
hprd_unique<- setdiff(HPRD, comp)

erichment.plot(hprd_unique, ultimate,"HPRD")
```
KEGG
```{r fig.width=12}
comp <- c(HPRD,PSP,WP, RXM, BG,SIG)
comp <- unique(comp)
kegg_unique<- setdiff(KEGG, comp)

erichment.plot(kegg_unique, ultimate,"KEGG")
```
WikiPathways
```{r fig.width=12}
comp <- c(HPRD,PSP,KEGG, RXM, BG, SIG)
comp <- unique(comp)
wp_unique<- setdiff(WP, comp)

erichment.plot(wp_unique, ultimate,"WikiPathways")
```
PhosphoSitePlus
```{r fig.width=12}
comp <- c(HPRD,KEGG,WP, RXM, BG, SIG)
comp <- unique(comp)
psp_unique<- setdiff(PSP, comp)
erichment.plot(psp_unique, ultimate,"PhosphoSitePlus")
```
BioGRID
```{r fig.width=12}
comp <- c(HPRD,KEGG,WP, RXM, PSP, SIG)
comp <- unique(comp)
bg_unique<- setdiff(BG, comp)

erichment.plot(bg_unique, ultimate,"BioGRID")
```
SIGNOR
```{r fig.width=12}
comp <- c(HPRD,KEGG,WP, RXM, PSP, BG)
comp <- unique(comp)
sig_unique<- setdiff(SIG, comp)

erichment.plot(sig_unique, ultimate,"SIGNOR")
```

